---
title: "2025_August_Unique_Type_Conversion"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Unique_Type_Conversion.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---




# Canker areas: Adding 0s for those without canker
If 'visible_canker' is NO then canker percentages (base/trunk/girdled) are 0.

The form using a Yes/No "Is there visible canker?" question. If the answer was "No" then the numeric questions (a_base_canker_area, a_trunk_canker_area, etc.) are skipped resulting in empty data entries. However, these entries do not represent the numeric value that should be there: they should be 0. 

## Adults
```{r}
cleaning <- after_health_assess %>% mutate(a_base_canker_area = if_else(a_visible_canker == "No", 0, a_base_canker_area))

cleaning <- cleaning %>% mutate(a_9_feet_canker_area = if_else(a_visible_canker == "No", 0, a_9_feet_canker_area))

cleaning <- cleaning %>% mutate(a_trunk_canker_area = if_else(a_visible_canker == "No", 0, a_trunk_canker_area))

cleaning <- cleaning %>% mutate(a_circum_girdled = if_else(a_visible_canker == "No", 0, a_circum_girdled))

cleaning_adults <- cleaning %>% filter(adult_or_seedling == "Adult")
original_adults <- after_health_assess %>% filter(adult_or_seedling == "Adult")
# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning_adults$site_name,
  plant_number = cleaning_adults$plant_number,
  original_a_visible_canker = original_adults$a_visible_canker,
  original_a_base_canker = original_adults$a_base_canker_area,
  clean_a_base_canker  = cleaning_adults$a_base_canker_area,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```

## Seedlings
```{r}
cleaning <- after_health_assess %>% mutate(s_base_canker_area = if_else(s_visible_canker == "No", 0, s_base_canker_area))

cleaning <- cleaning %>% mutate(s_stem_canker_area = if_else(s_visible_canker == "No", 0, s_stem_canker_area))

cleaning <- cleaning %>% mutate(s_circum_girdled = if_else(s_visible_canker == "No", 0, s_circum_girdled))

cleaning_seedlings <- cleaning %>% filter(adult_or_seedling == "Seedling")
original_seedlings <- after_health_assess %>% filter(adult_or_seedling == "Seedling")

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning_seedlings$site_name,
  plant_number = cleaning_seedlings$plant_number,
  original_a_visible_canker = original_seedlings$s_visible_canker,
  original_a_base_canker = original_seedlings$s_base_canker_area,
  clean_a_base_canker  = cleaning_seedlings$s_base_canker_area,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```