---
title: "2025_August_Unique_Type_Conversion"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Unique_Type_Conversion.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---

## Y/N to Yes/No
Consistent formating of boolean values to be "Yes" and "No"

### Epicormics 
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_base_epicormics = if_else(a_base_epicormics == "Y", "Yes", a_base_epicormics),
    a_base_epicormics = if_else(a_base_epicormics == "N", "No", a_base_epicormics),
    a_trunk_epicormics = if_else(a_trunk_epicormics == "Y", "Yes", a_trunk_epicormics),
    a_trunk_epicormics = if_else(a_trunk_epicormics == "N", "No", a_trunk_epicormics)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_base_ep = after_health_assess$a_base_epicormics,
  clean_base_ep  = cleaning$a_base_epicormics,
  
  original_trunk_ep = after_health_assess$a_trunk_epicormics,
  clean_trunk_ep    = cleaning$a_trunk_epicormics,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```

### Visible canker (adult & seedling)
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_visible_canker = if_else(a_visible_canker == "Y", "Yes", a_visible_canker),
    a_visible_canker = if_else(a_visible_canker == "N", "No", a_visible_canker),
    s_visible_canker = if_else(s_visible_canker == "Y", "Yes", s_visible_canker),
    s_visible_canker = if_else(s_visible_canker == "N", "No", s_visible_canker)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_a_vis_canker = after_health_assess$a_visible_canker,
  clean_a_vis_canker  = cleaning$a_visible_canker,
  
  original_s_vis_canker = after_health_assess$s_visible_canker,
  clean_s_vis_canker    = cleaning$s_visible_canker,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```


### Visible callous 
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_a_visible_callous = if_else(a_a_visible_callous == "Y", "Yes", a_a_visible_callous),
    a_a_visible_callous = if_else(a_a_visible_callous == "N", "No", a_a_visible_callous),
    s_a_visible_callous = if_else(s_a_visible_callous == "Y", "Yes", s_a_visible_callous),
    s_a_visible_callous = if_else(s_a_visible_callous == "N", "No", s_a_visible_callous)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_a_vis_callous = after_health_assess$a_a_visible_callous,
  clean_a_vis_callous  = cleaning$a_a_visible_callous,
  
  original_s_vis_canker = after_health_assess$s_a_visible_callous,
  clean_s_vis_callous   = cleaning$s_a_visible_callous,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```


## Canker areas: Adding 0s for those without canker
If 'visible_canker' is NO then canker percentages (base/trunk/girdled) are 0.

# Alive, adults
```{r}
cleaning <- after_health_assess %>% mutate(a_base_canker_area = if_else(a_visible_canker == "No", 0, a_base_canker_area))

cleaning <- after_health_assess %>% mutate(a_9_feet_canker_area = if_else(a_visible_canker == "No", 0, a_9_feet_canker_area))

cleaning <- after_health_assess %>% mutate(a_trunk_canker_area = if_else(a_visible_canker == "No", 0, a_trunk_canker_area))

cleaning <- after_health_assess %>% mutate(a_circum_girdled = if_else(a_visible_canker == "No", 0, a_circum_girdled))


# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_a_visible_canker = after_health_assess$a_visible_canker,
  original_a_base_canker = after_health_assess$a_base_canker_area,
  clean_a_base_canker  = cleaning$a_base_canker_area,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```


# Dead, adults
after_health_assess <- after_health_assess %>% mutate(a_d_base_canker_area = if_else(a_d_visible_callous == "No", 0, a_a_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_d_9_feet_canker_area = if_else(a_d_visible_callous == "No", 0, a_a_9_feet_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_d_trunk_canker_area = if_else(a_d_visible_callous == "No", 0, a_a_trunk_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_d_circum_girdled = if_else(a_d_visible_callous == "No", 0, a_a_circum_girdled))

after_health_assess <- after_health_assess %>% mutate(s_base_canker_area = if_else(s_a_visible_callous == "No", 0, s_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_stem_canker_area = if_else(s_a_visible_callous == "No", 0, s_trunk_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_circum_girdled = if_else(s_a_visible_callous == "No", 0, s_circum_girdled))