---
title: "2025_August_Unique_Type_Conversion"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Unique_Type_Conversion.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---

# Adjustments from the form structure
These are data changes required by the form structure. For instance, if "visible_canker" is "No" then canker percentages are skipped by the form and left as NA. Instead, these should all be 0. Thus, this space implements implicit updates to data values, due to the forms structure.

## Canker areas: Adding 0s for those without canker
If 'visible_canker' is NO then canker percentages (base/trunk/girdled) are 0.

The form using a Yes/No "Is there visible canker?" question. If the answer was "No" then the numeric questions (a_base_canker_area, a_trunk_canker_area, etc.) are skipped resulting in empty data entries. However, these entries do not represent the numeric value that should be there: they should be 0. 

### Adults
```{r}
cleaning <- after_health_assess %>% mutate(a_base_canker_area = if_else(a_visible_canker == "No", 0, a_base_canker_area))

cleaning <- cleaning %>% mutate(a_9_feet_canker_area = if_else(a_visible_canker == "No", 0, a_9_feet_canker_area))

cleaning <- cleaning %>% mutate(a_trunk_canker_area = if_else(a_visible_canker == "No", 0, a_trunk_canker_area))

cleaning <- cleaning %>% mutate(a_circum_girdled = if_else(a_visible_canker == "No", 0, a_circum_girdled))

cleaning_adults <- cleaning %>% filter(adult_or_seedling == "Adult")
original_adults <- after_health_assess %>% filter(adult_or_seedling == "Adult")
# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning_adults$site_name,
  plant_number = cleaning_adults$plant_number,
  original_a_visible_canker = original_adults$a_visible_canker,
  original_a_base_canker = original_adults$a_base_canker_area,
  clean_a_base_canker  = cleaning_adults$a_base_canker_area,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```

### Seedlings
```{r}
cleaning <- after_health_assess %>% mutate(s_base_canker_area = if_else(s_visible_canker == "No", 0, s_base_canker_area))

cleaning <- cleaning %>% mutate(s_stem_canker_area = if_else(s_visible_canker == "No", 0, s_stem_canker_area))

cleaning <- cleaning %>% mutate(s_circum_girdled = if_else(s_visible_canker == "No", 0, s_circum_girdled))

cleaning_seedlings <- cleaning %>% filter(adult_or_seedling == "Seedling")
original_seedlings <- after_health_assess %>% filter(adult_or_seedling == "Seedling")

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning_seedlings$site_name,
  plant_number = cleaning_seedlings$plant_number,
  original_a_visible_canker = original_seedlings$s_visible_canker,
  original_a_base_canker = original_seedlings$s_base_canker_area,
  clean_a_base_canker  = cleaning_seedlings$s_base_canker_area,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```

## Percent live canopy: Adding 0s for dead
If an individual is dead, we do not ask its percent live canopy in the form because it is implicitly 0. Here, I make the implicit, explicit.

### Adults
```{r}
cleaning <- after_health_assess

# Implement these changes
cleaning <- after_health_assess %>% mutate(a_percent_live_canopy = if_else(a_dead_or_alive == "Dead", 0, a_a_percent_live_canopy))  %>% select(-a_a_percent_live_canopy)


# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(adult_or_seedling == "Adult"))$site_name,
  plant_number = (cleaning %>% filter(adult_or_seedling == "Adult"))$plant_number,
  a_dead_or_alive = (cleaning %>% filter(adult_or_seedling == "Adult"))$a_dead_or_alive,
  original = (after_health_assess %>% filter(adult_or_seedling == "Adult"))$a_a_percent_live_canopy,
  clean  = (cleaning %>% filter(adult_or_seedling == "Adult"))$a_percent_live_canopy,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% mutate(a_percent_live_canopy = if_else(a_dead_or_alive == "Dead", 0, a_a_percent_live_canopy))  %>% select(-a_a_percent_live_canopy)

```

### Seedlings
```{r}
cleaning <- after_health_assess

# Implement change
cleaning <- after_health_assess %>% mutate(s_percent_live_canopy = if_else(s_dead_or_alive == "Dead", 0, s_a_percent_live_canopy))  %>% select(-s_a_percent_live_canopy)

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(adult_or_seedling == "Seedling"))$site_name,
  plant_number = (cleaning %>% filter(adult_or_seedling == "Seedling"))$plant_number,
  a_dead_or_alive = (cleaning %>% filter(adult_or_seedling == "Seedling"))$s_dead_or_alive,
  original = (after_health_assess %>% filter(adult_or_seedling == "Seedling"))$s_a_percent_live_canopy,
  clean  = (cleaning %>% filter(adult_or_seedling == "Seedling"))$s_percent_live_canopy,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% mutate(s_percent_live_canopy = if_else(s_dead_or_alive == "Dead", 0, s_a_percent_live_canopy))  %>% select(-s_a_percent_live_canopy)

```

## Purdue scores: Adding 5s if dead
```{r}
cleaning <- after_health_assess

# Implement these changes
cleaning <- after_health_assess %>% mutate(
  a_purdue_severity_canker = if_else(a_dead_or_alive == "Dead", "5", a_a_purdue_severity_canker),
  a_purdue_severity_canopy = if_else(a_dead_or_alive == "Dead", "5", a_a_purdue_severity_canopy)
        # Remove originals
    )  %>% select(-a_a_purdue_severity_canker, -a_a_purdue_severity_canopy)


# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(adult_or_seedling == "Adult"))$site_name,
  plant_number = (cleaning %>% filter(adult_or_seedling == "Adult"))$plant_number,
  a_dead_or_alive = (cleaning %>% filter(adult_or_seedling == "Adult"))$a_dead_or_alive,
  original_canker = (after_health_assess %>% filter(adult_or_seedling == "Adult"))$a_a_purdue_severity_canker,
  clean_canker  = (cleaning %>% filter(adult_or_seedling == "Adult"))$a_purdue_severity_canker,
  original_canopy = (after_health_assess %>% filter(adult_or_seedling == "Adult"))$a_a_purdue_severity_canopy,
  clean_canopy  = (cleaning %>% filter(adult_or_seedling == "Adult"))$a_purdue_severity_canopy,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% mutate(
  a_purdue_severity_canker = if_else(a_dead_or_alive == "Dead", "5", a_a_purdue_severity_canker),
  a_purdue_severity_canopy = if_else(a_dead_or_alive == "Dead", "5", a_a_purdue_severity_canopy)
        # Remove originals
    )  %>% select(-a_a_purdue_severity_canker, -a_a_purdue_severity_canopy)

```

# End