---
title: "2025_August_Categorical_Renamings"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Categorical_Renamings.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---

# Categorical renamings
These are the renamings of each categorical options. For instance, "Y"/"N"s to "Yes" and "No" or Purdue severity ratings from "1. Fewer than 3 active cankers that ..." to "1".

This also includes standardizing the NA responds across values like GPS ("N/a", "n/a", 999) all mean NA.


##  Plant Numbers
We decided that the plant numbers should rename a character value. This was because of some unique circumstances with some individuals requiring a character to distinguish. For instance, 206 and 206a, 67a and 67b, or NBL01.

However, many of them get entered with additional leading 0s like "001", "002", etc.

Here, I remove those leading 0s for consistency.
```{r}
cleaning <- after_health_assess %>%
  mutate(
    plant_number = sub("^0+", "", plant_number)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  original_plant_number = after_health_assess$plant_number,
  cleaned_plant_number = cleaning$plant_number,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```


## Site names
Correct capitalization inconsistencies with CPVT and Sugar River.

```{r}
# Correct capitalization inconsistencies with CPVT and Sugar River
after_health_assess <- after_health_assess %>%
  mutate(site_name = case_when(
    site_name == "Cpvt" ~ "CPVT",
    str_detect(site_name, "SR") ~ "SR", # Some of the SR's include the individual by accident, e.g. "SR-10"
    TRUE ~ site_name # Otherwise, keep name the same
  ))
```

## GPS - Numeric NA renamings
Presence of several different ways of saying an NA. Here I clarify them all to actual NA values.

```{r}
after_health_assess <- after_health_assess %>% mutate(
  gps_north = case_when(
    gps_north == 0 ~ NA,
    gps_north == "Na" ~ NA,
    gps_north == "N/A" ~ NA,
    gps_north == "X" ~ NA,
    gps_north == 999 ~ NA,
    TRUE ~ gps_north
  ),
  gps_west = case_when(
    gps_west == 0 ~ NA,
    gps_west == "Na" ~ NA,
    gps_west == "N/A" ~ NA,
    gps_west == "X" ~ NA,
    gps_north == 999 ~ NA,
    TRUE ~ gps_west
  )
)

```


## Renaming Y/N --> Yes/No

### Epicormics
We decided on consistent language for boolean True/False values to be set as Yes/No.
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_base_epicormics = if_else(a_base_epicormics == "Y", "Yes", a_base_epicormics),
    a_base_epicormics = if_else(a_base_epicormics == "N", "No", a_base_epicormics),
    a_trunk_epicormics = if_else(a_trunk_epicormics == "Y", "Yes", a_trunk_epicormics),
    a_trunk_epicormics = if_else(a_trunk_epicormics == "N", "No", a_trunk_epicormics)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_base_ep = after_health_assess$a_base_epicormics,
  clean_base_ep  = cleaning$a_base_epicormics,
  
  original_trunk_ep = after_health_assess$a_trunk_epicormics,
  clean_trunk_ep    = cleaning$a_trunk_epicormics,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```

### Visible canker (adult & seedling)
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_visible_canker = if_else(a_visible_canker == "Y", "Yes", a_visible_canker),
    a_visible_canker = if_else(a_visible_canker == "N", "No", a_visible_canker),
    s_visible_canker = if_else(s_visible_canker == "Y", "Yes", s_visible_canker),
    s_visible_canker = if_else(s_visible_canker == "N", "No", s_visible_canker)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_a_vis_canker = after_health_assess$a_visible_canker,
  clean_a_vis_canker  = cleaning$a_visible_canker,
  
  original_s_vis_canker = after_health_assess$s_visible_canker,
  clean_s_vis_canker    = cleaning$s_visible_canker,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```


### Visible callous 
```{r}
cleaning <- after_health_assess %>%
  mutate(
    # Seperating discretely to "Y"/"N" ensures that seedlings don't get assigned a epicormic value
    a_a_visible_callous = if_else(a_a_visible_callous == "Y", "Yes", a_a_visible_callous),
    a_a_visible_callous = if_else(a_a_visible_callous == "N", "No", a_a_visible_callous),
    s_a_visible_callous = if_else(s_a_visible_callous == "Y", "Yes", s_a_visible_callous),
    s_a_visible_callous = if_else(s_a_visible_callous == "N", "No", s_a_visible_callous)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  original_a_vis_callous = after_health_assess$a_a_visible_callous,
  clean_a_vis_callous  = cleaning$a_a_visible_callous,
  
  original_s_vis_canker = after_health_assess$s_a_visible_callous,
  clean_s_vis_callous   = cleaning$s_a_visible_callous,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))
```


## Seed estimate: Text to categorical
Before August+ we did not have consistent options, so people could type in whatever estimate they preferred. 

Based on the numeric values they inputted, I assign the same categories used in August+. In August+, the two options are "Few - less than 50 seeds on the tree" or "Lots - more than 50 seeds on the tree"
```{r}
# Add column to see what as.numeric does to the values 
cleaning <- after_health_assess %>% select(timestamp, site_name, plant_number, a_a_seed_estimate)

# Change to "Few (<50)" or "Lots (>50)"
cleaning <- cleaning %>% mutate(
  cleaned_seed_est = case_when(
    str_detect(a_a_seed_estimate, "Lots - more than 50 seeds on the tree") ~ "Lots (>50)",
    str_detect(a_a_seed_estimate, "Few - less than 50 seeds on the tree") ~ "Few (<50)"
  )
)

# Table of before and after for verification
datatable(
  cleaning %>% select(-timestamp),
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
cleaning <- cleaning %>% mutate(a_a_seed_estimate = cleaned_seed_est) %>% select(-cleaned_seed_est) 
after_health_assess <- after_health_assess %>% rows_update(cleaning, by = c("timestamp", "site_name", "plant_number"))
```


## Aspect
Consistency in naming
* Some entries had N/A & n/a --> NA 
* Other entries were fully typed out like: "North" --> N
```{r}
cleaning <- after_health_assess %>% select(timestamp, site_name, plant_number, aspect)

# To uppercase
cleaning <- cleaning %>% mutate(aspect = toupper(aspect))

cleaning <- cleaning %>% mutate(
  aspect_clean = case_when(
    aspect == "N/A" ~ NA,
    str_detect(aspect, "[0-9]") ~ NA, # For some slopes which were 0, aspect was also noted as 0/1, so if contains a number set to NA
    aspect == "WNW" ~ "NW",
    aspect == "North" ~ "N",
    aspect == "East" ~ "E",
    aspect == "South" ~ "S",
    aspect == "West" ~ "W",
    TRUE ~ aspect
  )
)

# Table of before and after for verification
datatable(
  cleaning %>% select(-timestamp),
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
cleaning <- cleaning %>% mutate(aspect = aspect_clean) %>% select(-aspect_clean) 
after_health_assess <- after_health_assess %>% rows_update(cleaning, by = c("timestamp", "site_name", "plant_number"))
```


## Upland/Riparian
Removes additional text description
```{r}
after_health_assess <- after_health_assess %>% mutate(
  upland_rip = recode(
    after_health_assess$upland_rip,
    "Upland: For our purposes, any land that is not riparian (even if not particularly hilly or even elevated)" = "Upland",
    "Riparian: On land immediately adjacent to rivers, streams, lakes, etc. or within flooding zone" = "Riparian",
  )
)
```

## Purdue severity ratings (Canker & Canopy)
Removes additional text description
```{r cleaning_purdue_1}
after_health_assess$a_a_purdue_severity_canker

# Note: If NA, leaves as NA
after_health_assess <- after_health_assess %>% mutate(
  a_a_purdue_severity_canker = recode(
    after_health_assess$a_a_purdue_severity_canker,
    "1. Fewer than 3 active cankers that are all smaller than 2-3 inches in length or diameter OR fewer than 3 inactive cankers." = "1",
    "2. More than 3 active cankers, OR 2-5 shallow (with no dead tissue) healed over with cracks less than 7 inches long." = "2",
    "3. More than 5 active OR inactive cankers cracked through the bark to the tissue below which have healed over, but you still see the level of damage." = "3",
    "4. Cankers occur all over the 10-foot area, with deep cracks and both active and inactive cankers." = "4",
    "5. Tree almost dead, mostly inactive cankers with deep cracks to dead tissue." = "5"
  )
)

# Note: If NA, leaves as NA
after_health_assess <- after_health_assess %>% mutate(
  a_a_purdue_severity_canopy = recode(
    after_health_assess$a_a_purdue_severity_canopy,
    "1. No apparent crown dieback." = "1",
    "2. Some but limited crown dieback." = "2",
    "3. Significant dieback in canopy, limited degree of new growth." = "3",
    "4. Major dieback in canopy, and limited or no new growth on terminal branches. Often segments of shredded or broken pieces of bark.  Epicormic branches present." = "4",
    "5. The canopy shows significant damage (large dead limbs, often broken limbs) also. Leaves only or mostly present as epicormic branches on trunk or base." = "5"
  )
)
```

## Signs of damage
Rename the options selected in the signs of damage to consistent forms
```{r}
# Select columns of interest for this correction
cleaning <- after_health_assess %>% select(timestamp, plant_number, site_name, adult_or_seedling, s_signs_damage)

# Signs of dmg only affects seedlings
cleaning <- cleaning %>% filter(adult_or_seedling == "Seedling") 

# Seperate rows by the comma-seperated values to correct naming directly 
cleaning <- cleaning %>% separate_rows(s_signs_damage, sep = ",\\s*") 

# Rename the individual responses 
cleaning <- cleaning %>% mutate(s_signs_damage = case_when(
  s_signs_damage == "Deer browse" ~ "Deer browse",
  s_signs_damage == "Deer rub" ~ "Deer rub",
  s_signs_damage == "Having been mowed in the past" ~ "Mowed",
  s_signs_damage == "Vole (nibbling on the bark)" ~ "Vole damage",
  s_signs_damage == "Rabbit browse" ~ "Rabbit browse",
  s_signs_damage == NA ~ "No damage",
  TRUE ~ "No damage"
))

# Recombine rows based on site, plant_number 
cleaning <- cleaning %>% group_by(timestamp) %>% mutate(s_signs_damage = paste(s_signs_damage, collapse = ", ")) %>% ungroup()

cleaning <- cleaning %>% distinct()

# Demonstrating changes 
original <- after_health_assess %>% select(timestamp, plant_number, site_name, adult_or_seedling, s_signs_damage)

cleaning <- original %>% rows_update(cleaning, by = c("timestamp", "site_name", "plant_number"))

comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  adult_or_seedling = cleaning$adult_or_seedling,
  original = original$s_signs_damage,
  cleaned = cleaning$s_signs_damage
)

comparison <- comparison %>%
  arrange(factor(adult_or_seedling, levels = c("Seedling", "Adult")))  

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by = c("timestamp", "site_name", "plant_number"))
cleaning <- after_health_assess # reset for testing
```

# End