
# Identification Variables
Here, I re-type and fix the primary problems with the most important identification variables prior to the data management which requires specifically referring to data entries.

Specifically, timestamp, plant numbers, site name and plant initials.


## Timestamp
Convert timestamp 8/15/2025 12:33:57 --> 2025-08-15 12:33:57
```{r}
after_health_assess <- after_health_assess %>% mutate(timestamp = mdy_hms(timestamp))
```

## Plant numbers
We decided that the plant numbers should rename a character value. This was because of some unique circumstances with some individuals requiring a character to distinguish. For instance, 206 and 206a, 67a and 67b, or NBL01.

However, many of them get entered with additional leading 0s like "001", "002", etc.

Here, I remove those leading 0s for consistency.
```{r}
cleaning <- after_health_assess %>%
  mutate(
    plant_number = sub("^0+", "", plant_number)
  )

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  original_plant_number = after_health_assess$plant_number,
  cleaned_plant_number = cleaning$plant_number,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```


## Site names
Correct capitalization inconsistencies with CPVT and Sugar River.

```{r}
# Correct capitalization inconsistencies with CPVT and Sugar River
after_health_assess <- after_health_assess %>%
  mutate(site_name = case_when(
    site_name == "Cpvt" ~ "CPVT",
    str_detect(site_name, "SR") ~ "SR", # Some of the SR's include the individual by accident, e.g. "SR-10"
    TRUE ~ site_name # Otherwise, keep name the same
  ))
```

## Plant initials
Standardizing the plant initials according to site.
```{r}
after_health_assess <- after_health_assess %>% mutate(plant_initials = case_when(
  site_name == "WCP" ~ "SH",
  site_name == "SR" ~ "SR",
  site_name == "ILM" ~ "BU",
  site_name == "CPVT" ~ "BU",
  TRUE ~ plant_initials
))
```

# End