
# Typing_without_introduced_NAs.R

Here I assign the correct data typings for the straightforward variables. 

These stand in contrast to variables "with_introduced_NAs". These variables that have data entry mistakes which are fixed during cleaning. Attempting to impose a typing before this cleaning would lose data. 

NA can accidentally be introduced when not all data points match the expected typing. For instance, if a character value is in a column of expected integers. If I try to impose all as integers, as I must to change the typing of the variable, we would lose that character value.  

The retyping is organized by which individual that variable applies to. 

## All individuals 
- GPS retyping in the final cleaning step due to necessary data cleaning
- Slope retyped in the final cleaning step due ||
- Height || 

```{r}
# Hybrid characters
after_health_assess <- after_health_assess %>% mutate(shape_terminal_bud = as.character(shape_terminal_bud))
after_health_assess <- after_health_assess %>% mutate(shape_lenticels = as.character(shape_lenticels))
after_health_assess <- after_health_assess %>% mutate(shape_leaf_scar = as.character(shape_leaf_scar))
after_health_assess <- after_health_assess %>% mutate(shape_hairs = as.character(shape_hairs))
```

## Adults

### All adults
```{r}
# Canopy 
after_health_assess <- after_health_assess %>% mutate(a_a_percent_live_canopy = as.integer(a_a_percent_live_canopy))

# Canker
after_health_assess <- after_health_assess %>% mutate(a_base_canker_area = as.integer(a_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_9_feet_canker_area = as.integer(a_9_feet_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_trunk_canker_area = as.integer(a_trunk_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_circum_girdled = as.integer(a_circum_girdled))
```

## Seedlings

### All seedlings
```{r}
# Densio
after_health_assess <- after_health_assess %>% mutate(s_densio_north = as.integer(s_densio_north))
after_health_assess <- after_health_assess %>% mutate(s_densio_east = as.integer(s_densio_east))
after_health_assess <- after_health_assess %>% mutate(s_densio_south = as.integer(s_densio_south))
after_health_assess <- after_health_assess %>% mutate(s_densio_west = as.integer(s_densio_west))

# Previous stems
after_health_assess <- after_health_assess %>% mutate(s_previous_stem_present = as.integer(s_previous_stem_present))

# Germ year
after_health_assess <- after_health_assess %>% mutate(s_germ_year_estimate = as.integer(s_germ_year_estimate))

# Canker
after_health_assess <- after_health_assess %>% mutate(s_base_canker_area = as.integer(s_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_stem_canker_area = as.integer(s_stem_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_circum_girdled = as.integer(s_circum_girdled))
```

### Alive seedlings
```{r}
# Canopy
after_health_assess <- after_health_assess %>% mutate(s_a_percent_live_canopy = as.integer(s_a_percent_live_canopy))
```

# End