
# Typing_without_introduced_NAs.R

Here I assign the correct data typings for the straightforward variables. 

These stand in contrast to variables "with_introduced_NAs". These variables that have data entry mistakes which are fixed during cleaning. Attempting to impose a typing before this cleaning would lose data. 

NA can accidently be introduced when not all data points match the expected typing. For instance, if a character value is in a column of expected integers. If I try to impose all as integers, as I must to change the typing of the variable, we would lose that character value.  

The retyping is organized by which individual that variable applies to. 

# All individuals 
- GPS retyping in the final cleaning step due to necessary data cleaning
- Slope retyped in the final cleaning step due ||

# Adults

## All adults
```{r}
after_health_assess <- after_health_assess %>% mutate(a_base_canker_area = as.integer(a_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_9_feet_canker_area = as.integer(a_9_feet_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_trunk_canker_area = as.integer(a_trunk_canker_area))
after_health_assess <- after_health_assess %>% mutate(a_circum_girdled = as.integer(a_circum_girdled))
```

# Seedlings
```{r}
after_health_assess <- after_health_assess %>% mutate(s_base_canker_area = as.integer(s_base_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_stem_canker_area = as.integer(s_stem_canker_area))
after_health_assess <- after_health_assess %>% mutate(s_circum_girdled = as.integer(s_circum_girdled))
```
