---
title: "2025_June_July_Identification_Variables"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Identification_Variables.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(DT)
library(dplyr)
library(styler)
library(lubridate) # For working with dates and times

original <- june_health_assess

# # Imports all June/July 2025 data
# june_health_assess <- read_csv("../data/Raw_Datasets/2025_June_July_Butternut_Health_Assessment_Form_Responses.csv")
# 
# source("../data_cleaning/2025_June_July_1_Preparing_Columns.R", local=TRUE)
# source("../data_cleaning/2025_June_July_2_Type_Conversions.R", local=TRUE)
# source("../data_cleaning/2025_June_July_3_Removing_Test_Entries.R", local=TRUE)
# source("../data_cleaning/2025_June_July_4_Processing_Dead_Adults.R", local=TRUE)
# 
# exists("june_health_assess")
```

# Identification 

Note: Order of cleaning operations DOES matter as site and plant_number must be corrected prior to all other corrections. These variables serve to identify each individual entry ensuring that corrections are matched directly to the correct entry. 

## Site names

Override all the individuals within this timeframe to have a site_name of WCP; correct capitalization inconsistencies with CPVT and Sugar River.

```{r}
# Site name is WCP for all trees up until July 1st
last_wcp_time <- ymd_hms("2025-07-01 10:57:46")

june_health_assess <- june_health_assess %>%
  mutate(site_name = ifelse(timestamp < last_wcp_time, "WCP", site_name))

# Correct capitalization inconsistencies with CPVT and Sugar River
june_health_assess <- june_health_assess %>%
  mutate(site_name = case_when(
    site_name == "Cpvt" ~ "CPVT",
    site_name == "Sugar river" ~ "SR",
    site_name == "Sugar River" ~ "SR",
    TRUE ~ site_name # Otherwise, keep name the same
  ))

```

## Plant numbers to numeric

Parsing the plant numbers as numbers rather than strings.

```{r}
# Remove 'SH' or 'SH ' from the beginning of plant numbers 
june_health_assess$plant_number <- sub("^SH\\s*", "", june_health_assess$plant_number)

cleaning <- june_health_assess
cleaning$plant_number <- parse_number(cleaning$plant_number)

# Demonstrating changes
comparison <- tibble(
  original_plant_number = june_health_assess$plant_number,
  clean_plant_number = cleaning$plant_number
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying changes
june_health_assess <- june_health_assess %>% mutate(plant_number = as.integer(parse_number(plant_number)))
```

## Adult and Seedling Labels
Currently, the original question was "Is this individual a seedling?" so the inputs are all Yes or No.

### a. Converting Yes/No to S/A
First, convert all Yes responses into "Seedling" and Nos to "Adult". Then we will need to manually set the values for the individuals from the first day of assessment where we didn't have that question added yet. 
```{r}
june_health_assess <- june_health_assess %>% mutate(adult_or_seedling = case_when(
  adult_or_seedling == "Yes" ~ "Seedling",
  adult_or_seedling == "No" ~ "Adult"
))
```

### b. Manually setting missing labels
Manually setting the adult/seedling value for the individuals done prior to that question being added.

For the individuals from June 5th and 6th, if the densiometer is not empty label as a seedling, otherwise adult.
```{r}
june_6th = ymd_hms("2025-06-6 23:59:59") # "this is black walnut!"

june_health_assess <- june_health_assess %>%
  mutate(adult_or_seedling = if_else(
    timestamp < june_6th,
    # If TRUE (before June 6th), update labels by densiometer:
    case_when(
      is.na(densio_north) ~ "Adult",
      TRUE ~ "Seedling" 
    ),
    # If FALSE (June 6th or later), keep whatever was already there
    adult_or_seedling
  ))

```

## Plant initials
Standardizing the plant initials according to site.
```{r}
june_health_assess <- june_health_assess %>% mutate(plant_initials = case_when(
  site_name == "WCP" ~ "SH",
  site_name == "SR" ~ "SR",
  site_name == "ILM" ~ "BU",
  site_name == "CPVT" ~ "BU",
  TRUE ~ plant_initials
))

```



```{r}
# Remove extra df created for demonstration
rm(list = setdiff(ls(), "june_health_assess"))

# 1. Create a data frame of the column names and their precise internal type (e.g., double, integer)
data_types_table <- data.frame(
  Variable_Name = names(june_health_assess),
  Specific_Data_Type = sapply(june_health_assess, typeof),
  row.names = NULL
)

# 2. View it as a beautifully formatted table right in your console
knitr::kable(data_types_table, format = "markdown")

```
