---
title: "2025_June_July_Identification_Variables"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = "2025_June_July_Identification_Variables.html",
      output_dir = "C:/Users/helmerhj/Documents/GitHub/2026_Butternut_Health_Data_Management/communication/rpub"
    )
  })
---

# Identification 
Here, I re-type and fix the primary problems with the most important identification variables prior to the data management which requires specifically referring to data entries.

Specifically, timestamp, plant numbers, site name, and plant initials.

Note: Order of cleaning operations DOES matter as site and plant_number must be corrected prior to all other corrections. These variables serve to identify each individual entry ensuring that corrections are matched directly to the correct entry. 

## Timestamp
Convert timestamp 8/15/2025 12:33:57 --> 2025-08-15 12:33:57

```{r}
june_health_assess <- june_health_assess %>% mutate(timestamp = mdy_hms(timestamp))
```

## Plant numbers -- remove excess text

Parsing the plant numbers as numbers rather than strings.

```{r}
# Remove 'SH' or 'SH ' from the beginning of plant numbers 
june_health_assess$plant_number <- sub("^SH\\s*", "", june_health_assess$plant_number)

cleaning <- june_health_assess
cleaning$plant_number <- parse_number(cleaning$plant_number)

# Demonstrating changes
comparison <- tibble(
  site = june_health_assess$site_name,
  original = june_health_assess$plant_number,
  clean = as.character(cleaning$plant_number)
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying changes
june_health_assess <- june_health_assess %>% mutate(plant_number = as.character(parse_number(plant_number)))
```

## Site names

Override all the individuals within this timeframe to have a site_name of WCP; correct capitalization inconsistencies with CPVT and Sugar River.

```{r}
# Site name is WCP for all trees up until July 1st
last_wcp_time <- ymd_hms("2025-07-01 10:57:46")

june_health_assess <- june_health_assess %>%
  mutate(site_name = ifelse(timestamp < last_wcp_time, "WCP", site_name))

# Correct capitalization inconsistencies with CPVT and Sugar River
june_health_assess <- june_health_assess %>%
  mutate(site_name = case_when(
    site_name == "Cpvt" ~ "CPVT",
    site_name == "Sugar river" ~ "SR",
    site_name == "Sugar River" ~ "SR",
    TRUE ~ site_name # Otherwise, keep name the same
  ))

```

### b. Manually setting missing labels
Manually setting the adult/seedling value for the individuals done prior to that question being added.

For the individuals from June 5th and 6th, if the densiometer is not empty label as a seedling, otherwise adult.
```{r}
june_6th = ymd_hms("2025-06-6 23:59:59") # "this is black walnut!"

june_health_assess <- june_health_assess %>%
  mutate(adult_or_seedling = if_else(
    timestamp < june_6th,
    # If TRUE (before June 6th), update labels by densiometer:
    case_when(
      is.na(densio_north) ~ "Adult",
      TRUE ~ "Seedling" 
    ),
    # If FALSE (June 6th or later), keep whatever was already there
    adult_or_seedling
  ))

```

## Plant initials
Standardizing the plant initials according to site.
```{r}
june_health_assess <- june_health_assess %>% mutate(plant_initials = case_when(
  site_name == "WCP" ~ "SH",
  site_name == "SR" ~ "SR",
  site_name == "ILM" ~ "BU",
  site_name == "CPVT" ~ "BU",
  TRUE ~ plant_initials
))

```

# End