
# Typing 
Here I assign the correct data typings for the straightforward variables. 

These stand in contrast to variables "with_introduced_NAs". These variables that have data entry mistakes which are fixed during cleaning. Attempting to impose a typing before this cleaning would lose data. 

NA can accidentally be introduced when not all data points match the expected typing. For instance, if a character value is in a column of expected integers. If I try to impose all as integers, as I must to change the typing of the variable, we would lose that character value.  

The retyping is organized by which individual that variable applies to. 

## Seedling variables
```{r}
june_health_assess <- june_health_assess %>% mutate(shape_terminal_bud = as.character(shape_terminal_bud))
june_health_assess <- june_health_assess %>% mutate(shape_lenticels = as.character(shape_lenticels))
june_health_assess <- june_health_assess %>% mutate(shape_leaf_scar = as.character(shape_leaf_scar))
june_health_assess <- june_health_assess %>% mutate(shape_hairs = as.character(shape_hairs))
```

## ****Canker areas to integer 
NOTE: Right now, "Less than 10, but more than 0" just reads in as 10
  
Note: Using 'parse_number' first is very important. Parse_number retrieves the number from
     within a string, removing excess characters, e.g. parse_number("70%") = 70.

Then, as.integer turns it into an integer, but if you used only as.integer any excess characters would lead to parsing the whole value as NA, e.g. as.integer("70%") = NA.

```{r}
# Trunk canker area
june_health_assess$trunk_canker_area <- as.integer(parse_number(june_health_assess$trunk_canker_area))

# Circumference girdled
june_health_assess$circum_girdled_canker <- as.integer(parse_number(june_health_assess$circum_girdled_canker))

# Base canker area
june_health_assess$base_canker_area <- as.integer(parse_number(june_health_assess$base_canker_area))

# Percent live canopy
june_health_assess$percent_live_canopy <- as.integer(parse_number(june_health_assess$percent_live_canopy))
```

# End