
# June / July Dead Individuals
We did not record dead/alive during June/July. 

Here I search for dead/alive in the notes and this variable update accordingly.
Further comfirmation will happen during photo review in QC. 

## WCP - known dead
- 38; "DEAD seedling - no   sign of canker. Appears to have been death by shade"
- 2: "DEAD tree. Canopy class is based on structure, not live canopy. Callous is also based on whether cankers look like they have ever been calloused."
- 50: "Stem is dead. Need to add a category for "seedling died" Unsure if 0 canker or 100% canker. Could have been killed by canker or the fungus is from after it died."
- 3 (two entries one): We did this assessment based on dead adult tree. Epicormic health characterization will be entered as its own form.
- 10: "Dead"

```{r}
# List of dead individuals 
dead_WCP = c(38, 2, 50, 3, 10)

# Isolate WCP dead individuals 
cleaning <- june_health_assess %>% mutate(
  a_dead_or_alive = case_when(
    # If a WCP individual within the dead list AND an adult
    (site_name == "WCP" & plant_number %in% dead_WCP & adult_or_seedling == "Adult") ~ "Dead",
    # All other WCP (for June/July) adults are alive
    (site_name == "WCP" & !(plant_number %in% dead_WCP) & adult_or_seedling == "Adult") ~ "Alive",
    TRUE ~ a_dead_or_alive
  ),

  s_dead_or_alive = case_when(
    # If a WCP individual within the dead list AND an seedling
    (site_name == "WCP" & plant_number %in% dead_WCP & adult_or_seedling == "Seedling") ~ "Dead",
    # All other WCP (for June/July) seedling are alive
    (site_name == "WCP" & !(plant_number %in% dead_WCP) & adult_or_seedling == "Seedling") ~ "Alive",
    TRUE ~ s_dead_or_alive
  )
)

# Show before and after 
# Demonstrating changes
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  adult_or_seedling = cleaning$adult_or_seedling,
  og_adult_status  = june_health_assess$a_dead_or_alive,
  clean_a = cleaning$a_dead_or_alive,
  og_seedling_status = june_health_assess$s_dead_or_alive,
  clean_s = cleaning$s_dead_or_alive
)

# Table of before and after for verification
datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
june_health_assess <- june_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))

```

## ILM - known dead
- 211: Dead; Dale estimated that the main trunk died 10 years ago. *Epicormic may be graft-able. 
- 214: DEAD, in super dense shady cedar stand
- 219: DEAD tree! - snag, broken in half. Last photo taken was not written down.
- 245: DEAD!
- 343: DEAD. lesion on base. 5.5mm base diameter
- 344: DEAD, less grassy competition than nearby seedlings. 5.6mm base diameter
- 345: 4.2mm basal diameter DEAD...
- 346: DEAD, triple stem - all dead..
- 347: 7.5mm basal diameter       DEAD
- 348: This is a re-sprout from a stem that is dead (which was equally tall)
- 354: DEAD...
- 376: DEAD...
- 378: DEAD...
```{r}
dead_ILM = c(211, 214, 219, 245, 343, 344, 345, 346, 347, 348, 354, 376, 378)

cleaning <- june_health_assess %>% mutate(
  # Adults
  a_dead_or_alive = case_when(
    # If a WCP individual within the dead list AND an adult 
    (site_name == "ILM" & plant_number %in% dead_ILM & adult_or_seedling == "Adult") ~ "Dead",
    # All other WCP (for June/July) adults are alive
    (site_name == "ILM" &!(plant_number %in% dead_ILM) & adult_or_seedling == "Adult" ~ "Alive"),
    TRUE ~ a_dead_or_alive
  ),
  
  s_dead_or_alive = case_when(
    # If a WCP individual within the dead list AND an seedling 
    (site_name == "ILM" & plant_number %in% dead_ILM & adult_or_seedling == "Seedling") ~ "Dead",
    # All other WCP (for June/July) seedling are alive
    (site_name == "ILM" & !(plant_number %in% dead_ILM) & adult_or_seedling == "Seedling" ~ "Alive"),
    TRUE ~ s_dead_or_alive
  )
)

# Show before and after 
# Demonstrating changes
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  adult_or_seedling = cleaning$adult_or_seedling,
  og_adult_status  = june_health_assess$a_dead_or_alive,
  clean_a = cleaning$a_dead_or_alive,
  og_seedling_status = june_health_assess$s_dead_or_alive,
  clean_s = cleaning$s_dead_or_alive
)

# Table of before and after for verification
datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
june_health_assess <- june_health_assess %>% rows_update(cleaning, by=c("timestamp", "plant_number", "site_name"))

```


## All other individuals, assume alive
```{r}

# If not already assigned, assume alive.
cleaning <- june_health_assess %>% mutate(
  # Adults
  a_dead_or_alive = case_when(
    is.na(a_dead_or_alive) ~ "Alive",
    TRUE ~ a_dead_or_alive
  ),
  
  # Seedlings
  s_dead_or_alive = case_when(
    is.na(s_dead_or_alive) ~ "Alive",
    TRUE ~ s_dead_or_alive
  ),
)


# Show before and after 
# Demonstrating changes
comparison <- tibble(
  site_name = june_health_assess$site_name,
  plant_number = june_health_assess$plant_number,
  adult_or_seedling = june_health_assess$adult_or_seedling,
  a_dead_or_alive  = june_health_assess$a_dead_or_alive,
  s_dead_or_alive = june_health_assess$s_dead_or_alive
)

# Table of before and after for verification
datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

```

# End