
# Data Entry Corrections 

## All Individuals

### GPS
1. 2025-09-12 09:40:47 ILM 586 "--73.329" --> remove "-73.329"
```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-09-12 09:40:47")

# ILM 586 "--73.329" --> remove "-73.329"
row_index <- which(cleaning$timestamp == individuals_timestamp)
cleaning$gps_west[row_index] <- -73.329

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og_gps_west = (after_health_assess %>% filter(timestamp == individuals_timestamp))$gps_west,
  cleaned = (cleaning %>% filter(timestamp == individuals_timestamp))$gps_west
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))

```

### Slope
One individual had an input slope of "Yes (no measurement tool with me)"
The photos for this individual were unavailable, so --after talking with Emma--
we decided to go with NA.

#### June/July 
Sugar River #3 (July individual)
- NA for slope
- NA for aspect 
```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-07-03 10:05:45")

# 2025-09-15T08:38:10Z, SR 3, slope="Yes (no measurement tool with me)"
row_index <- which(cleaning$timestamp == individuals_timestamp)
cleaning$slope[row_index] <- NA
cleaning$aspect[row_index] <- NA

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og_slope = (after_health_assess %>% filter(timestamp == individuals_timestamp))$slope,
  cleaned_slope = (cleaning %>% filter(timestamp == individuals_timestamp))$slope,
  og_aspect = (after_health_assess %>% filter(timestamp == individuals_timestamp))$aspect,
  cleaned_aspect = (cleaning %>% filter(timestamp == individuals_timestamp))$aspect
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

#### August 
1. 2025-09-15T08:38:10Z, ILM	#500	slope="o", should be 0

```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-09-15T08:38:10Z")

# 2025-09-15T08:38:10Z, ILM 500, slope="o"
row_index <- which(cleaning$timestamp == individuals_timestamp)
cleaning$slope[row_index] <- 0

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og_slope = (after_health_assess %>% filter(timestamp == individuals_timestamp))$slope,
  cleaned_slope = (cleaning %>% filter(timestamp == individuals_timestamp))$slope
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

2. 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA, should be flipped
```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-09-19T13:48:15Z")

# 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA
row_index <- which(cleaning$timestamp ==individuals_timestamp)
cleaning$slope[row_index] <- 0
cleaning$aspect[row_index] <- NA

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og_slope = (after_health_assess %>% filter(timestamp == individuals_timestamp))$slope,
  cleaned_slope = (cleaning %>% filter(timestamp == individuals_timestamp))$slope,
  og_aspect = (after_health_assess %>% filter(timestamp == individuals_timestamp))$aspect,
  cleaned_aspect = (cleaning %>% filter(timestamp == individuals_timestamp))$aspect
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

### Plant heights
1. 2025-09-15 08:38:10, ILM	#3	plant_height_ft = E
```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-09-15 08:38:10")

# 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA
row_index <- which(cleaning$timestamp ==individuals_timestamp)
cleaning$plant_height_ft[row_index] <- 0

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og_slope = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_height_ft,
  cleaned_slope = (cleaning %>% filter(timestamp == individuals_timestamp))$plant_height_ft
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

## Adults

### DBH_cm 
Some entries were written with two DBHs, like so:
1. 2025-11-11T16:17:47Z	HPCA	14	DBH="17.9, 14.0"
2. 2026-06-05T15:08:18Z	WCP	10	DBH="34.5, 44"
3. 2025-08-27T10:31:09Z	CPVT	29	DBH="43.6, 30.7"

Here I go through each, review the notes and update accordingly. 
For those that were split tree trunks, I corrected the height to the LARGER of the two DBHs, then updated the additional note to include this update. 

1. 2025-11-11T16:17:47Z	HPCA	14	DBH="17.9, 14.0"
Data enterer: Catherine dell'Olio
Note said: "NA", alive adult
Messaged Catherine (https://docs.google.com/spreadsheets/d/1SC2aXJ5x1a-RNEMqzuhe5M4UJ23KxcVWEW7tb96s0CI/edit?disco=AAACAQ_hBKc) to ask what the meaning

```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2025-11-11T16:17:47Z")

# 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA
row_index <- which(cleaning$timestamp ==individuals_timestamp)
cleaning$a_DBH_cm[row_index] <-17.9
cleaning$additional_notes <- paste0(cleaning$additional_notes, ", two-stemmed split trunk with DBHs 17.9, 14.0. The greater was chosen during data cleaning - Hank Helmers.")

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og = (after_health_assess %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  cleaned = (cleaning %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  notes = (cleaning %>% filter(timestamp == individuals_timestamp))$additional_notes
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

2. 2026-06-05T15:08:18Z	WCP	10	DBH="34.5, 44"
Data enterer: Catherine dell'Olio
Note said: "NA", dead adult

```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2026-06-05T15:08:18Z")

# 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA
row_index <- which(cleaning$timestamp ==individuals_timestamp)
cleaning$a_DBH_cm[row_index] <-44.0
cleaning$additional_notes <- paste0(cleaning$additional_notes, ", two-stemmed split trunk with DBHs 34.5, 44. The greater was chosen during data cleaning - Hank Helmers.")

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og = (after_health_assess %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  cleaned = (cleaning %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  notes = (cleaning %>% filter(timestamp == individuals_timestamp))$additional_notes
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

3. 2025-08-27T10:31:09Z	CPVT	29	DBH="43.6, 30.7"
Data enterer: Kylie Roth
Note said: "low split with 2 main branches", alive, adult

```{r}
cleaning <- after_health_assess

individuals_timestamp = ymd_hms("2026-06-05T15:08:18Z")

# 2025-09-19T13:48:15Z, ILM	#364, aspect = 0, slope = NA
row_index <- which(cleaning$timestamp ==individuals_timestamp)
cleaning$a_DBH_cm[row_index] <-44.0
cleaning$additional_notes <- paste0(cleaning$additional_notes, ", two-stemmed split trunk with DBHs 34.5, 44. The greater was chosen during data cleaning - Hank Helmers.")

# Demonstrating changes 
comparison <- tibble(
  site_name = (cleaning %>% filter(timestamp == individuals_timestamp))$site_name,
  plant_number = (after_health_assess %>% filter(timestamp == individuals_timestamp))$plant_number,
  og = (after_health_assess %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  cleaned = (cleaning %>% filter(timestamp == individuals_timestamp))$a_DBH_cm,
  notes = (cleaning %>% filter(timestamp == individuals_timestamp))$additional_notes
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% rows_update(cleaning, by=c("timestamp", "site_name"))
```

