# Typing with introduced NAs
This cleaning step corrects the typing of variables that had mistyped entries leading to introduced NAs when attempting to impose a type on it.

For instance, one GPS entry said "--73.223". If uncorrected during the cleaning step, running as.double("--73.223") would result in gps_west=NA, effectively erasing that data point.

Instead, these mistyped errors are corrected in "Data_Entry_Corrections.Rmd" and then re-typed here. 

## All individuals 

### GPS: Retype to double
One GPS entry said "--73.223".
```{r}
cleaning <- after_health_assess

# Apply retyping
cleaning <- cleaning %>% mutate(gps_west = as.double(gps_west)) 
cleaning <- cleaning %>% mutate(gps_north = as.double(gps_north))

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  og_gps_north = after_health_assess$gps_north,
  cleaned_gps_north = cleaning$gps_north,
  og_gps_west = after_health_assess$gps_west,
  cleaned_gps_west = cleaning$gps_west,
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% mutate(gps_west = as.double(gps_west)) 
after_health_assess <- after_health_assess %>% mutate(gps_north = as.double(gps_north))

```

### Slope: Type to integer


## Height
One height entry said "E", a likely misentry from the aspect question right before the height question in the form.

## Plant height
```{r}
cleaning <- after_health_assess

# Apply retyping
cleaning <- after_health_assess %>% mutate(plant_height_ft = as.double(plant_height_ft))

# Demonstrating changes 
comparison <- tibble(
  site_name = cleaning$site_name,
  plant_number = cleaning$plant_number,
  og_height = after_health_assess$plant_height_ft,
  cleaned_height = cleaning$plant_height_ft
)

datatable(
  comparison,
  options = list(
    pageLength = 10,
    scrollY = "400px",
    scrollX = TRUE
  ),
  class = "stripe hover row-border order-column" # forces light theme
)

# Applying the changes
after_health_assess <- after_health_assess %>% mutate(plant_height_ft = as.integer(plant_height_ft))
```
